---
title: "Lab 9: Baby Names"
subtitle: "Regression"
author: "Your name"
format: 
  html:
    self-contained: true
    code-tools: true
    toc: true
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
#| include: false
library(broom)
library(tidyverse)
```

[Download .qmd starter file](Lab9-babynames.qmd)

[Download `StateNames_E.csv`](baby-names-data/StateNames_E.csv)

Our data set in this lab contains baby names that begin with "E" and their popularity over time. For each year, and for each name with at least 50 recorded babies born, we are given the counts of how many babies had that name.

::: callout-note
If you wish to download the data set for your name, you can find the names for ALL 50 states, in separate data sets organized by first letter of the name [here](https://github.com/earobinson95/stat331-calpoly/tree/master/lab-assignments/lab9/baby-names-data).
:::

Let's take a look at how the name "Emily" has changed over time. As my name begins with "E", you should load in the `StateNames_E.csv` data set.

```{r}
#| include: false
babies_e <- read.csv(
  here::here("labs",
             "Lab 9",
             "StateNames_E.csv")
  )
```

## Explore the Number of Emilys

**1. _Create a Game Plan & Code!_ -- Make a summary table of the number of babies named "Emily" for each state and the sex of the baby.**

  + **Specifically, each state should be its own row and each sex should have its own column.**
  + **Additionally, if there were no babies born with the name "Emily" for that combination of state & sex the table should have a value of 0 (not an `NA`).** 
  
::: callout-warning 
To be clear, there should be **one** row per state, which means you need to `summarize()` all the years for each state.
:::

::: callout-caution
The dataset has a column titled `Gender`, which contains two values `"F"` and `"M"`, representing "Female" and "Male". The sex someone was assigned at birth is different from their gender identity [(definitions)](https://www.plannedparenthood.org/learn/gender-identity/sex-gender-identity). Thus, this variable should be renamed to `Sex` or `Sex at Birth`.

You should be using only **one** {dplyr} pipeline!
:::

![Game Plan](https://imgur.com/a/zqulKPx)

```{r}
num_emily <- babies_e |> 
  filter(Name == "Emily") |>
  rename(Sex = Gender) |> 
  pivot_wider(names_from = Sex,
              values_from = Count) |> 
  select(State, M, `F`, Year) |> 
  mutate(M = replace_na(M, 0)) |> 
  group_by(State) |>
  summarize(Males = sum(M),
            Females = sum(`F`))
  num_emily
```

**2. You should have seen in the table above that "Emily" is a name given overwhelmingly to babies assigned "female" at birth. Create a new data set to include only babies named "Emily" assigned `F`emale at birth. Summarize the data so there is one observation (the total number of Emilys born) for each year.**

::: callout-note
+ Remember you wanted to rename the variable "Gender"
+ Save this data set so you can use it in the rest of the lab questions.
+ No output needed for this question, only code
:::

```{r}
female_emilys <- babies_e |> 
  filter(Name == "Emily",
         Gender == "F") |>
  rename(Sex = Gender) |> 
  select(Name, Year, State, Count) |> 
  group_by(Year) |> 
  summarize(`Number of Emilys` = sum(Count))
female_emilys
```

**3. Using the new summarized data set from above, make a visualization showing how the popularity of the name "Emily" for babies assigned female at birth has changed over the years.**

```{r}
female_emilys |> 
  ggplot(aes(x=Year,
             y=`Number of Emilys`)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Popularity of Females Assigned at Birth Named Emily",
       x = "Year",
       subtitle = "Count",
       y = NULL)
```

## Modeling the Number of Emilys

**4. Using the new summarized data set from above, create a linear model with the year as the explanatory variable, and the number of Emilys as the response.**

```{r}
lm_emily <- lm(`Number of Emilys` ~ Year,
               data = female_emilys)
tidy(lm_emily)
```

**5. Visualize the linear regression.**

```{r}
female_emilys |> 
  ggplot(aes(x = Year,
             y = `Number of Emilys`)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Relationship between Year and Frequency of Emilys",
       x     = "Year",
       y     = NULL,
       subtitle = "Count") +
  theme_bw()
```

**6. Write out the estimated regression equation.**

```{r}
tidy(lm_emily)
```


**My Regression Equation:**

$$\hat{Num Emilys} = 1963731.36 - 968.93 * Year $$

Formatting taken from [here](https://rpruim.github.io/s341/S19/from-class/MathinRmd.html)

**7. Plot the _residuals_ of the model versus the Year. Comment on the residuals - do you see any patterns?**

::: callout-tip
The `augment()` function from the **broom** package may be helpful!
:::

```{r}
augment(lm_emily) |> 
  ggplot(aes(x = Year,
             y = `.resid`)) +
  geom_point() +
  geom_hline(yintercept = 0,
             color = "red") +
  labs(title = "Residuals Plot",
       x = "Year",
       y = NULL,
       subtitle = "Residual") +
  theme_bw()
```

It appears that the residuals plot takes on a wave-like pattern, indicating a linear model may not be the best fit.

**8. What do you conclude from this model? Is my name not cool anymore?**

I believe the name Emily is decreasing in popularity yearly. There must be cooler names.


## Spelling by State

In middle school I was **so** upset with my parents for not naming me "Emmaleigh". Past my pre-teen rebellion, I'm happy with my name and am glad when baristas spell it "Emily".

**9. Narrow the original E name data set down to only female-assigned babies with alternative spellings -- "Emilee", "Emilie", "Emmaleigh", or "Emmalee".**

```{r}
alt_names <- babies_e |> 
  rename(Sex = Gender) |> 
  filter(Name %in% c("Emilee", "Emilie", "Emmaleigh", "Emmalee"),
         Sex == "F")
head(alt_names)
```

**10. Make a plot comparing the popularity (total count) of these alternate spellings over time.**

```{r}
alt_names |> 
  pivot_wider(names_from = Name,
              values_from = Count) |> 
    mutate(across(.cols = Emilee:Emmaleigh,
                .fns = ~replace_na(., 0))) |>
  select(State, Emilee, Emilie, Emmaleigh, Emmalee, Year) |> 
  group_by(Year) |> 
  summarize(Emilee = sum(Emilee),
            Emilie = sum(Emilie),
            Emmaleigh = sum(Emmaleigh),
            Emmalee = sum(Emmalee)) |> 
  pivot_longer(cols = Emilee:Emmalee,
               values_to = "Count",
               names_to = "Spelling") |> 
  ggplot(aes(x = Year,
             y = Count,
             group = Spelling,
             color = Spelling)) +
  geom_line() +
  labs(title = "Popularity of Emily 'Variants' Over Time",
       x = "Year",
       y = NULL,
       subtitle = "Count") +
  theme_bw()
```

**11. The spelling of Emmaleigh is the least common of the four alternatives. However, the spelling picked up popularity in the Southeast (North Carolina, Georgia, and Florida). Create a table to display the number of babies born with each alternative spelling in each of these these three Southeast states, as well as in California, during 2012.**

::: callout-warning
Specifically, each of the four alternative spellings should be in their own column and each of the four states should have their own row.
:::

::: callout-caution
You should be using only **one** {dplyr} pipeline!
:::

```{r}
alt_names |> 
  filter(State %in% c("NC",
                      "GA",
                      "FL",
                      "CA"),
         Year == 2012) |> 
  pivot_wider(names_from = Name,
              values_from = Count) |> 
  group_by(State) |> 
  summarize(Emilee = sum(Emilee),
            Emilie = sum(Emilie),
            Emmalee = sum(Emmalee),
            Emmaleigh = sum(Emmaleigh))
  
```


**12. Create a table that converts your total counts to overall percents. That is, what was the percent breakdown between the four alternative spellings in each of the four states?**

::: callout-caution
You should be using only **one** {dplyr} pipeline!
:::

```{r}
# Code Q12
```